### 2025-10-17 write/response_to_reviews_GA.tex
updated analysis et al for revision

\documentclass[12pt]{article}
% Language and font
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
% Bibliography with APA style
\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber]{biblatex}
\DeclareLanguageMapping{american}{american-apa}
\addbibresource{sample.bib}
% Layout and formatting
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\setstretch{1.0}
\usepackage{parskip} % disables indentation and adds vertical space between paragraphs
\setlength{\parindent}{0pt}
% Math and symbols
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{gensymb}
% Graphics and floats
\usepackage{graphicx}
\usepackage{float}
\usepackage{caption}
% Optional: nicer font
\usepackage{lmodern}
% Turn off section numbers
\setcounter{secnumdepth}{0}

\title{\begin{large}\textbf{Response to Reviews} \\ \vspace{.1in}
Is there a Criterion in Criterial Learning? \\ Insights from Studying Feedback Delays \\ \vspace{.1in}
\textit{Journal of Experimental Psychology: Learning, Memory, and Cognition} \\ \vspace{-.1in}
Manuscript number: XLM-2024-0436 \end{large}}
\author{
    Matthew J. Crossley, 
    Benjamin O. Pelzer,
    F. Gregory Ashby
}
\date{}

\begin{document}

\maketitle

We would like to thank the editor and reviewers for their
excellent suggestions and thoughtful comments. We worked
very hard to respond to all of these. As a result, we are
confident that the revised manuscript is significantly
stronger than our original submission. Please note that we
changed the title of the manuscript to more directly convey
the main point of our paper.

Our specific response to each reviewer comment is as
follows.

\section{Reviewer 1}
We were pleased that Reviewer 1 ``really enjoyed reading the
manuscript'' and ``highly recommend(ed) this paper to be
published after addressing a number of mostly minor
concerns.''

\subsection{Reviewer comment 1}
1. Exploration of model space: When reading the description
of the models it was not entirely clear to me whether the
three presented models explore the space of possible models
to a sufficient degree or whether there might be some models
missing from consideration that could change the overall
message.

\subsubsection{Author response}
This is an excellent and important point. In response, we
greatly expanded the set of models we explored, and we also
now classify these models more systematically. The previous
version examined 3 models, whereas the revision examines 12.
These 12 models are partitioned into two classes: one that
assumes a criterion is learned and one that assumes SR
learning with no criterion. The criterial-learning class
includes 8 different models that were created by factorially
combining three different binary assumptions: perceptual
drift (yes or no), criterial drift (yes or no), and
delay-sensitive updating (yes or no). The SR learning class
includes four models created by factorially combining two
binary assumptions: perceptual drift (yes or no) and
delay-sensitive feedback (yes or no).

\subsection{Reviewer comment 2}
2. Simulation results figures: There were some aspects of
the results figures for the simulation that I found
difficult to understand:

(a) Even after reading the full paper and thinking about it
for a while, I do not understand what panels B are supposed
to show. Does this only show the subset of the parameter
space in which one of the two relevant patterns was
observed? If not, shouldn't there be data points everywhere?

(b) I do not understand why in Figure 1, for example, panel
C shows some cases in which the predicted qualitative
pattern is either ``control impaired'' or ``other'', but there
is no corresponding bar in panel A.

(c) The bars in Figure 2A are smaller than in Figures 1A and
3A. Why?

\subsubsection{Author response}
With the new approach to model exploration, we have heavily
revised the simulation results figures.  We hope that the
new figures are clearer and address the reviewer's concerns.

\subsection{Reviewer comment 3}
3. Calling the simulation Experiment 1: I found it extremely
confusing that the simulation using parameter-space
partitioning was called ``Experiment 1''. If I read
``Experiment'' I expect to see some empirical data obtained
from participants and not solely a simulation. I suggest
renaming this section to something clearer. For example
``Simulation Study'' or ``Parameter-Space Partitioning
Across
Models''.

\subsubsection{Author response}
After thinking about this a bit, we agree that referring to
the computational modeling analyses as Experiment 1 was a
mistake. First, we agree with the reviewer that this
nomenclature is confusing, but we also came to realize that
the equations in this section might dissuade empirically
oriented readers from reading about our empirical results.
As a result, the revised manuscript renames the section
describing the computational modeling ``New Computational
Models of Criterial Learning'' and this more technical
section has been moved to later in the manuscript -- after
the two empirical experiments are described.

\subsection{Reviewer comment 4}
4. Outliers in Exp. 2:

- While I think the argument for removing outliers in Exp. 2
is fine, the actual criterion defining an outlier is not
actually given (which feels a bit ironic in a paper about
criteria).

\subsubsection{Author response}
Good point. We now say ``for each participant, we used the
best-fitting two component model to compute the likelihood
that their performance was best described by each component.
Participants whose data were estimated to most likely belong
to the low-performance distribution were classified as
outliers and excluded from further analysis.''

\subsection{Reviewer comment 5}
4. Outliers in Exp. 2:

- How do the results look with these outliers included? Does
the difference between condition vanish? Adding a footnote
or so should be enough.

- The fact that some participants did not even learn a
single criterion feels very surprising to me. Is this common
in such experiments? Please add a sentence embedding this
finding within your general experience running such
experiments.

\subsubsection{Author response}
When the outliers are included in the analysis, the
difference between conditions does indeed vanish. We have
added the following footnote: ``Including the excluded
outliers in these analyses eliminates the difference between
conditions.''

In our experience, it is not unusual for a small subset of
participants to perform very poorly in experiments of this
kind. This typically reflects a lack of engagement or
misunderstanding of the task instructions rather than an
inability to perform the task itself. We now state exactly
that in the revision.

\subsection{Reviewer comment 6}
Minor points:

- Does Eq. 13 show results for trial n or n + 1? The text
and equation are inconsistent.

\subsubsection{Author response}
We corrected this typo.

\subsection{Reviewer comment 7}
- The description regarding the randomisation of Experiment
2 on page 12 (paragraph before ``Procedure'') seems to be
inconsistent with Figure 4. Is the dimension picked at
random for each trial or always the same for certain problem
numbers (as implied by Figure 4)?

\subsubsection{Author response}
We agree that the description of the design of this
experiment (now Experiment 1) needed improvement. We revised
the paragraph Reviewer 1 refers to in this comment as
follows.

``Each participant practiced each of 14 one-dimensional
category-learning tasks, or problems, until they responded
correctly on 9 of the 10 previous trials, at which point the
problem changed. The 14 different category structures are
described in Figure 1. The relevant dimension was bar
thickness in problems 1 – 7 and bar angle in problems 8 –
14. Each problem included stimuli in two distinct clusters
that varied over a restricted range of the relevant
dimension. Critically, the optimal criterion value varied
from problem-to-problem with respect to its position within
the stimulus range. For example, for some problems the
optimal criterion was below the midpoint of the range and
for other problems it was above the midpoint.''

\subsection{Reviewer comment 8}
- In figure 4, the top says Problems ``1 thru 13'', but only
shows 7 problems.

\subsubsection{Author response}
We corrected this typo.

\vspace{.2in}

\section{Reviewer 2}

We were pleased that Reviewer 2 believed our  ``manuscript
reports a sophisticated approach to an important problem''
and that the reviewer ``especially like(d) the use of
mathematical models.''

\subsection{Reviewer comment 1}
I was concerned that the models seem to have multiple
varying features. It might be more useful to isolate
specific processes while holding other model elements
constant.

\subsubsection{Author response}
This is the same as Reviewer 1's first comment. As described
there, we strongly agree with this concern, and as a result,
we greatly expanded the number of models we examined, and we
reclassified these according to the fundamental assumption
of whether or not they assumed a criterion value is
specifically learned and stored trial-by-trial in memory.

\subsection{Reviewer comment 2}
I think it makes more sense to have a prediction simulation
section and Experiments 1 and 2 rather than to consider the
simulation the first experiment.

\subsection{Author response}
This is the same as Reviewer 1's third comment. As noted
there, we agree and as a result, we now refer to the two
empirical experiments as Experiment 1 and 2, and the
computational modeling section has been renamed to ``New
Computational Models of Criterial Learning'' and moved to
after the two experiments.

\subsection{Reviewer comment 3}
The data in Figure 6 can likely be fit much more closely by
an ExGaussian distribution than either the 1 or 2 component
Gaussian mixtures that are shown.

\subsubsection{Author response}
We thank the reviewer for this suggestion. We now estimate
both a single-component truncated ExGaussian and a
two-component truncated ExGaussian mixture via MLE with
proper normalization on the tasks bounded support $[9,
512]$, and compared them via AIC, BIC, and a parametric
bootstrap likelihood-ratio test (LRT).  The results from
moving from the mixture of Gaussians used in our orginal
submission to the truncated ExGaussian mixture used in the
revision does not change the qualitative pattern of our
results in any way. However, we agree that the ExGaussian
and ExGaussian mixture better capture the skewed and bounded
distribution of our trials-to-criterion data so we have
adopted it in the revision.

\subsection{Reviewer comment 4}
Replication of the criterial learning results from the
currently-labeled Experiment 2 seems like a good idea given
the fairly low sample sizes per condition.

\subsubsection{Author response}
We agree with the reviewer that a replication would be
valuable, but we believe the data and the expanded modeling
analysis currently presented represent a significant
contribution to the field.

\subsection{Reviewer comment 5}
Did stimuli remain on the screen across the feedback delay?
I don't think so. If not, what happens if the stimulus
reappears at the time of feedback? I could imagine this
would eliminate or dramatically reduce the negative effect
of feedback delay. If so, is that consistent with the
conclusions? That is, would the models that predict a
deleterious effect of feedback delay also predict that it
should be eliminated with ``reminder'' stimuli at the time
of feedback?

\subsubsection{Author response}
As shown in Figures 2 and 5, the stimulus was presented up
until a response was made. At this time the stimulus was
removed and replaced by a noise mask until feedback was
delivered. The question of how our results might have
changed if the stimulus were to reappear at the time of
feedback is both interesting and insightful. In the very
first study that investigated the effects of feedback delays
on category learning (Maddox, Ashby, \& Bohil, 2003,
JEP:LM$\&$C) we found that a short 2.5 s feedback delay
impaired information-integration category learning, but not
rule-based learning -- but only if a noise mask was
presented during the time between response and feedback. The
published 2003 article doesn't explore this issue, but we
hypothesized at the time that in the absence of the mask,
participants might be able to use eidetic imagery to keep
the relevant synapses active during the delay period. As a
result, we agree with the reviewer's prediction that
re-presenting the stimulus at the time of feedback would
eliminate the impaired learning that we observed. This
interesting point seems tangential to our main point,
however, and as a result, we did not address this point in
the revision.

\subsection{Reviewer comment 6}
For the currently-labeled Experiment 3, reporting
fail-to-reject results for t-tests is not a valid method for
establishing that feedback delay has no effect. This is
especially a concern given the low participant count per
condition and the fact that empirical effects go in the
``right'' direction (lower performance for delayed
feedback).

\subsubsection{Author response}
This is a completely valid point. As in almost all
psychological experiments, our Experiment 2 lacks the
statistical power to make strong inferences about the null
results we obtained there. We now explicitly acknowledge
this point in paragraph 2 of the section entitled
``Discussion of Experiments 1 and 2.'' Also, rather than try
to make a statistical argument in support of our null
results, we now use a converging operations approach by
appealing to the literature. The new relevant paragraph is
as follows.

``Whereas the results of Experiment 1 seem clear, the
Experiment 2 results are weaker since they suggest a null
result. Although insufficient power makes it impossible to
rule out the possibility that Experiment 2 missed some small
effect of feedback delay, the null results found there are
consistent with a wide variety of other evidence. In
particular, a variety of other studies have reported that
feedback delays of up to 10 s do not significantly slow
one-dimensional rule-based learning (Dunn et al., 2012; Ell
et al., 2009; Maddox et al., 2003; Maddox \& Ing, 2005;
Worthy et al., 2013). In addition, many studies have
reported evidence that one-dimensional rule-based learning
recruits working memory and executive attention, but not
procedural learning (for reviews, see, e.g., Ashby \&
Valentin, 2017; Ashby et al., 2020). So the results of
Experiment 2 are consistent with all of these studies. On
the other hand, several of these feedback-delay studies did
not use binary-valued stimulus dimensions, so some criterial
learning was required. If so, then why did they all report
no effect of feedback delay? It is important to note that
criterial learning was not the focus of any of these
studies, and as a result, in all of these previous studies
the optimal criterion was always set exactly midway between
the category prototypes, which makes criterial learning
trivial. For example, under these conditions, criterial
learning might not even require feedback. The unsupervised
category-learning experiments reported by Ashby et al.
(1999) provide strong support for this because all of their
rule-based participants learned the correct criterion (which
was midway between the category means), even though the task
was completely unsupervised.''

\subsection{Reviewer comment 7}
I recommend running a bigger replication study that combines
both tasks to demonstrate the claimed interaction (feedback
delay impairs performance only for the criterion learning
task). The replication study should be preregistered with a
sample size justification based on the interaction effect
defined by the across-experiment comparison in the current
data. Adding a ``stimulus reminder at feedback'' condition
to this new study might be helpful in clarifying the
mechanisms producing the delay effect.

\subsubsection{Author response}
We agree that a replication / extension study would
strengthen our manuscript -- as it would for almost all
empirical papers. We chose not to pursue this suggestion,
however.  Even so, we believe we did significantly
strengthen our manuscript by greatly expanding the class of
computational models that we investigated, and by better
situating our results within the existing literature.

\section*{Action Editor}

\subsection{AE comment 1}
The final paragraph of the introduction summarizes the
experiments. This seems more appropriate in the abstract,
given that no information about the experiments are
presented in the abstract.

\subsubsection{Author response}
Thanks for this suggestion. The revised abstract now has
more detail about the experiments.

\subsection{AE comment 2}
Equation 1 needs some explanation, as this implies that the
decision Rn is made immediately on stimulus presentation. In
other words, Rn is not an outcome of some time-extended
cognitive process. Am I misunderstanding equation 1?

\subsubsection{Author response}
This is a valid and insightful point. Equation 1 does assume
that the response is made immediately upon stimulus
presentation. Because we are not trying to account for
response time, we decided to not explicitly model the
time-extended cognitive processes that govern evidence
accumulation. We now state this explicitly in the revision.
Specifically, immediately after equation 1, we added the
following text:

``Note that this equation indicates that the response is
made immediately upon stimulus presentation. We therefore
are not explicitly modeling the time-extended cognitive
processes that govern evidence accumulation and the
corresponding response time.''

\subsection{AE comment 3}
What was the reason behind the choice to have equation 2
only be sensitive to negative feedback?

\subsubsection{Author response}
The rationale behind the choice to have equation 2 only be
sensitive to negative feedback is now stated explicitly in
the first two sentences of the paragraph that includes
equation 2. Those sentences read as follows.

``If positive feedback is received, then the value of the
criterion remains unchanged for the next trial (except for
drift – see equation 3). The rationale here is that if the
response is correct, then the observer has effectively
gained zero information about how their criterion should be
modified.''

\subsection{AE comment 4}
How is ``optimal'' defined in the current case, given that
response time does not factor in the update rule?

\subsubsection{Author response}
By ``optimal'', we mean the value of the criterion that
maximizes accuracy. We now say this explicitly in the
revision. This is now explained in the last sentence of the
paragraph that includes equation 2, which reads as follows.

``This updating rule will gradually converge on the optimal
criterion value (i.e., the value of the criterion that
maximizes accuracy).''

\subsection{AE comment 5}
The use of parameter space partitioning is sensible given
the many differences among the three models. However, PSP
operates on non-stochastic models. In addition, the way you
have approached PSP is not how the original PSP is
implemented (i.e., using MCMC). This should be mentioned
explicitly to avoid confusion.

\subsubsection{Author response}
Thanks for raising this important point. We have revised the
manuscript to make this point clear by adding the following
text  to the beginning of the third paragraph of the section
entitled ``Simulation Results.''

``Classical PSP is defined for non-stochastic models and is
often implemented by exploring the parameter space via a
Markov chain Monte Carlo algorithm. In contrast, we used
brute force search over a grid of plausible parameter
values. Because our models are stochastic, at each sampled
set of parameter values we generated multiple simulations
and classified the results into one of a small set of
qualitative outcome patterns.''

\subsection{AE comment 6}
The three-dimensional plots in figures 1B, 2B, and 3B do not
come out right on 2D. You could improve these by having 2D
projections on the planes. Better still, three boxplots
would be more appropriate as there is no additional
information by using 3D plots.

\subsubsection{Author response}
This was a good suggestion. We revised the two figures (8
and 9) that describe the simulation results accordingly --
i.e., both figures are 2D and include boxplots.

\subsection{AE comment 7}
Although not a focus in this manuscript, I wondered whether
the models are able to account for finer-grained data
patterns such as complete RT distributions and
speed-accuracy tradeoff functions.

\subsubsection{Author response}
As described in our response to AE comment 2, the models
make no attempt to account for response time (RT). The
models could all be generalized to make RT predictions, but
any such generalization would require adding extra
assumptions about the time course of the decision-making
process. The focus of our research was on \textit{what} is
updated trial-by-trial (i.e., a criterion vs SR
associations), rather than on the dynamics of the updating
process. For this reason, we chose a theoretically minimal
landscape. Specifically, we were concerned that adding any
extra assumptions that were not directly related to our
primary goal would weaken the inferences possible from our
model simulations. For example, it would be difficult to
rule out the possibility that any qualitative result we
established was a consequence of one of these added
assumptions, rather than because of a difference in what was
being updated on each trial. Even so, we agree that
extending the models proposed here to make RT predictions
could be a fruitful avenue for future work.

\subsection{AE comment 8}
The experiments are numbered 2 and 3, instead of 1 and 2.
The models were following the methods of experiment 1, which
is missing. Experiment 1 is also mentioned in the discussion
separately with experiment 2. It looks like experiment 1 has
been deleted from this manuscript, but the model simulations
and discussion still refer to it.

\subsubsection{Author response}
We thank the reviewer for catching this confusion. No study
was deleted from the manuscript. Rather, whether or not the
computational modeling section should be considered an
``experiment'' was a point of oscillation when writing. This
comment, along with comment 3 of Reviewer 1 and comment 2 of
Reviewer 2 helped us see that referring to the modeling
section as Experiment 1 was a mistake. As explained in our
responses to both of those comments, the revision refers to
the two empirical experiments as Experiments 1 and 2. These
are now presented first, and then following by a modeling
section entitled ``New Computational Models of Criterial
Learning.''

\subsection{AE comment 9}
In figure 6, it is clear that the participants form two
subcohorts. However, the use of a Gaussian seems odd given
that there can not be any negative values for
trials-to-criterion. Given that the authors model this
frequency histogram to exclude participants, more
appropriate distributions, such as Poisson should be used.

\subsubsection{Author response}
This is a good point that is similar to comment 3 of
Reviewer 2. Rather than use a mixture model based on the
Poisson distribution, we took the advice of Reviewer 2 and
used the ExGaussian as our base distribution. Specifically,
as described in our response to comment 3 of Reviewer 2, we
now fit the trials-to-criterion data collapsed across
conditions and participants with both a single-component
truncated ExGaussian and a two-component truncated
ExGaussian mixture via MLE with proper normalization on the
tasks bounded support $[9,512]$.

\subsection{AE comment 10}
Given the experimental results, what is the added benefit of
the three models? Experiment 2 is the critical experiment
showing that 3.5-0.5 leads to more trials-to-criterion than
0.5-3.5 or 0.5-0.5. The models do not provide any detailed
understanding of mechanisms and they are not fitted to
actual data. The general message from the experiments is
that any updating rule should be sensitive to feedback
delay. The models do not go beyond this.

\subsubsection{Author response}
The experiments show that feedback delay impairs criterial
learning, but they do not speak to the question of
underlying mechanism. The modeling is critical to this
question. Specifically, the computational analysis attempts
to establish constraints on underlying mechanisms that are
compatible with the empirical results. We believe that PSP
is a better approach, given this goal, than the more
traditional approach of fitting the models to the observed
data. This is now explained in the following new paragraph
(paragraph 2 of the ``Simulation Results'' section):

``Our approach was to generate predictions for each of the
three experimental conditions across a wide range of
parameter values. We then used parameter-space partitioning
(PSP) to evaluate the performance of each model (Pitt et
al., 2006). Traditional computational modeling identifies
parameter values that allow the model to provide the best
possible fit to the observed data. This traditional approach
therefore treats the single observed data set being fit as a
gold standard. Our goal instead was to investigate the range
of possible predictions of each model. We were interested in
questions such as: can a model predict any possible outcome
or is it constrained to always predict, for example, that
feedback delays impair learning? PSP was designed to answer
such questions.''

Furthermore, we believe this PSP analysis produced some
surprising and important results. First, it showed that
SR-learning accounts are compatible with our results -- even
if they assume that the amount of synaptic updating that
occurs trial-by-trial is unaffected by feedback delay.
Second, the PSP analysis showed that, in contrast,
stored-criterion accounts of Experiments 1 and 2 must assume
that feedback delays impair the criterion-updating process
-- a result that greatly constrains the type of memory that
could be used to store and update the criterion. For
example, it rules out working memory because many previous
results show that working memory is immune to short 3 s
delays. We believe these inferences would be impossible
without our computational modeling approach. In this way,
the models contribute the most important piece: they pin
down the process-level mechanisms consistent with the
experiments.

\end{document}