### 2025-10-21 write/cover_letter_revision.tex
bug fix

\documentclass[12pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\setstretch{1.0}
\signature{Matthew J. Crossley \\ on behalf of all authors}
\date{\today}

\begin{document}

\begin{letter}{Editorial Office \\ 
\textit{Journal of Experimental Psychology: Learning, Memory, and Cognition}}

\opening{Dear Colleagues,}

Please find enclosed our detailed response to the reviewers’
and action editor’s comments on our manuscript entitled
\textit{“Is there a Criterion in Criterial Learning?
Insights from Studying Feedback Delays”} (Manuscript
\#XLM-2024-0436).

We have carefully considered all feedback and made extensive
revisions throughout the paper. These include a substantial
expansion of the model space, clearer presentation of
simulation results, reorganization of manuscript sections,
and several textual clarifications. We believe these changes
have significantly strengthened the manuscript and improved
its clarity and theoretical contribution.

We sincerely thank you and the reviewers for the insightful
comments and constructive suggestions that guided these
improvements.

\closing{Sincerely,}

\end{letter}

\end{document}